

gray triggerfish Video Index Using ZINB
========

This analysis generates a gray triggerfish video index using a zero-inflated Negative Binomial and Poisson model structure as implemented in the R package pscl.

```{r setup, echo=FALSE,message=FALSE, warning=FALSE, error=FALSE}
rm(list=ls(all=TRUE)) 
graphics.off()
#windows(record=T)
#setwd ("\\\\CCFHR-S-1534090\\popdyn1\\SEDAR\\SEDAR 41\\SEFISIndices\\S41RGIT")
library(MASS)
library(doBy)
library(statmod)
library(Hmisc)
library(pscl)
library(lmtest)
library(stargazer)

```



```{r ReadData, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
# Read in gray triggerfish Data
# This is the updated file from Nate
raw=read.csv("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\SEDAR41_GT_VI\\data\\graytriggerfish_SEDAR41_2015v2.csv")
names(raw)
summary(raw)

```



```{r mungingData, echo=FALSE, message=FALSE, warning=FALSE}
gt<-raw
# Full data subsetting: 
gt <- gt[gt$Station_Type!="Recon",]                # remove recon stations
gt <- gt[gt$A.Video.Readable == "Yes",]         # remove invalid videos
gt <- subset(gt, gt$Start_Depth > 0)                  # remove NA in depth
gt <- subset(gt, gt$Start_Depth < 100)              # remove < 100 m deep
gt <- subset(gt, gt$LastOfTemp > 0)                 # remove blank water temps
gt <- subset(gt, gt$Turbidity != "Unknown")    # remove unknown turbidity values
gt <- subset(gt, gt$No.Readable.Frames ==41)    # remove zero readable frames
gt <- subset(gt, gt$Substrate !='Unknown')    # remove Substrate with unknowns
gt <- subset(gt, gt$Biotic_Density !='Unknown')    # remove biotic density with unknowns
gt <- subset(gt, gt$Year > 2010)    # remove 2010 samples as per workshop rec

names(gt)
length(gt$MeanCount)


#make the Substrate_Cat
temp=rep(NA,length(gt$Substrate))
temp[gt$Substrate=='Unknown']='Unknown'
gt$Substrate[gt$Substrate=='Unknown']=NA
gt$Substrate=as.numeric(levels(gt$Substrate)[gt$Substrate])
temp[gt$Substrate==0]='None'
temp[gt$Substrate>0 & gt$Substrate<10]='Low'
temp[gt$Substrate>=10 & gt$Substrate<40]='Moderate'
temp[gt$Substrate>=40]='High'
gt=cbind(gt,temp)
names(gt)=c(names(gt)[-dim(gt)[2]],'Substrate_Cat')

#make the Biotic_Density_Cat
temp=rep(NA,length(gt$Biotic_Density))
temp[gt$Biotic_Density=='Unknown']='Unknown'
gt$Biotic_Density[gt$Biotic_Density=='Unknown']=NA
gt$Biotic_Density=as.numeric(levels(gt$Biotic_Density)[gt$Biotic_Density])
temp[gt$Biotic_Density==0]='None'
temp[gt$Biotic_Density>0 & gt$Biotic_Density<10]='Low'
temp[gt$Biotic_Density>=10 & gt$Biotic_Density<40]='Moderate'
temp[gt$Biotic_Density>=40]='High'
gt=cbind(gt,temp)
names(gt)=c(names(gt)[-dim(gt)[2]],'Biotic_Density_Cat')

#rename startTime

names(gt)
#Eliminate unnecessary columns
dat=subset(gt,select=c(SumCount,No.Readable.Frames,Year,Turbidity,Current_Direction,
                       Current_Magnitude,Substrate_Cat,Relief,Size,Biotic_Density_Cat,
                       Biotic_Type,Biotic_Height,Start_Depth,Julian,Start_Latitude,
                       LastOfTemp))

orgnames=names(dat)



# sumcount=round(dat$MeanCount*gt$No.Readable.Frames,0)
# dat$MeanCount=sumcount



#rename to short names
names(dat)=c('SumCount','frames','y','wc','cd','cm','sc','sr','ss','bd','bt','bh','d','t','lat','temp')
head(dat)

#replace NA in the cpue with 0
dat$SumCount[is.na(dat$SumCount)]=0

# Table 1 data
stargazer(dat, type='text')
dat.2011<-subset(dat, dat$y==2011)
stargazer(dat.2011, type='text')
dat.2012<-subset(dat, dat$y==2012)
stargazer(dat.2012, type='text')
dat.2013<-subset(dat, dat$y==2013)
stargazer(dat.2013, type='text')
dat.2014<-subset(dat, dat$y==2014)
stargazer(dat.2013, type='text')
summaryBy(SumCount~y,data=dat,FUN=length)

#rescale frames
#dat$frames=dat$frames/41
#dat$frames=log(dat$frames)
```



```{r EDA, echo=FALSE, error=FALSE, warning=FALSE, fig.align='center'}
#now categorize the continuous variables

# Figure for original continus variable distributions
par(mfrow=c(2,2))
## DEPTH
hist(dat$d,breaks=seq(10,105,by=5),
     main="Depth",
     xlab="Depth (m)")
## LATITUDE
hist(dat$lat,breaks=seq(27,36,by=0.25),
     main="Latitude",
     xlab="Latitude (degrees)")
## DAY OF YEAR (t)
hist(dat$t,breaks=seq(110,305,by=5),
     main="Day of the Year",
     xlab="Julian Day")
## WATER TEMP (temp)
hist(dat$temp,breaks=seq(12.25,29.50,by=0.25),
     main="Temperature",
     xlab="Bottom Temperature (C)")


## DEPTH
hist(dat$d,breaks=seq(10,110,by=5))
summary(dat$d)

#temp=cut(dat$d,breaks=c(14,25,41,52,115),labels=FALSE)#;temp;table(temp)
#temp=cut(dat$d,2,breaks=c(0,as.numeric(summary(dat$d))[-c(1,4)]),labels=FALSE);temp;table(temp)
temp=cut(dat$d,2,breaks=quantile(dat$d),labels=FALSE)#;temp;table(temp)
dat$d=temp
summary(dat$d)

## LATITUDE
hist(dat$lat,breaks=seq(27,36,by=0.25))
#hist(dat$lat[dat$y=='2010'],breaks=seq(27,36,by=0.25))
hist(dat$lat[dat$y=='2011'],breaks=seq(27,36,by=0.25))
hist(dat$lat[dat$y=='2012'],breaks=seq(27,36,by=0.25))
hist(dat$lat[dat$y=='2013'],breaks=seq(27,36,by=0.25))
summary(dat$lat)
#temp=cut(dat$lat,breaks=c(27,29.75,31.25,32.75,34,35.25),labels=FALSE)#;temp;table(temp)
#temp=cut(dat$lat,2,breaks=c(0,as.numeric(summary(dat$lat))[-c(1,4)]),labels=FALSE);temp;table(temp)
#temp=cut(dat$lat,2,breaks=quantile(dat$lat),labels=FALSE)#;temp;table(temp)
temp=cut(dat$lat,2,breaks=quantile(dat$lat),labels=FALSE)#;temp;table(temp)
dat$lat=temp
summary(dat$lat)
hist(dat$lat)

## DAY OF YEAR (t)
hist(dat$t,breaks=seq(110,305,by=5))
summary(dat$t)
#temp=cut(dat$t,breaks=c(##,##,##,##,##,##),labels=FALSE)#;temp;table(temp)
#temp=cut(dat$t,2,breaks=c(0,as.numeric(summary(dat$t))[-c(1,4)]),labels=FALSE);temp;table(temp)
##temp=cut(dat$t,2,breaks=quantile(dat$t),labels=FALSE)#;temp;table(temp)
temp=cut(dat$t,2,breaks=quantile(dat$t,probs=seq(0,1,0.125)),labels=FALSE)#;temp;table(temp)
dat$t=temp
summary(dat$t)
hist(dat$t)

## WATER TEMP (temp)
summary(dat$temp)
hist(dat$temp,breaks=seq(12.25,29.50,by=0.25))

#temp=cut(dat$t,breaks=c(##,##,##,##,##,##),labels=FALSE)#;temp;table(temp)
#temp=cut(dat$temp,2,breaks=c(0,as.numeric(summary(dat$temp))[-c(1,4)]),labels=FALSE);temp;table(temp)
temp=cut(dat$temp,2,breaks=quantile(dat$temp),labels=FALSE)#;temp;table(temp)
dat$temp=temp
summary(dat$temp)
hist(dat$temp)

# ## TIME OF DAY (tod)
# hist(dat$tod,breaks=seq(0.4,0.95,by=0.025))
# summary(dat$tod)
# #temp=cut(dat$tod,breaks=c(##,##,##,##,##,##),labels=FALSE)#;temp;table(temp)
# #temp=cut(dat$tod,2,breaks=c(0,as.numeric(summary(dat$tod))[-c(1,4)]),labels=FALSE);temp;table(temp)
# temp=cut(dat$tod,2,breaks=quantile(dat$tod),labels=FALSE)#;temp;table(temp)
# dat$tod=temp


#Factorize variables
dat$y=factor(dat$y)
dat$d=factor(dat$d)
dat$lat=factor(dat$lat)
dat$t=factor(dat$t)
dat$temp=factor(dat$temp)
# dat$tod=factor(dat$tod)

#get rid of unused factors for dat$wc
dat$wc=factor(dat$wc)
dat$sc=factor(dat$sc)
dat$bd=factor(dat$bd)

plot.design(SumCount~y + wc + cd,data=dat)
plot.design(SumCount~ sc + bd ,data=dat)
plot.design(SumCount~d + t + lat + temp ,data=dat)

```

**So this bit evaluates the ZIP versus the ZINB **

The ZINB is clearly preferred in the likelihood ratio test and fits the data better.  But you can see that neither model fit the data particularly well.


```{r ZIP-ZINB, echo=T,message=T, warning=T, cache=TRUE}

zipform=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp )
#zipform=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y  )
#zipform=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |1.  )
zipmod=zeroinfl(zipform,  dist = "poisson", link = "logit",data=dat);summary(zipmod)

nbform=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp )
#nbform=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y )
#nbform=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |1. )
nbmod=zeroinfl(nbform,  dist = "negbin", link = "logit",data=dat);summary(nbmod)

PvNP<-lrtest(zipmod,nbmod)

stargazer(zipmod,nbmod, type="text", 
          title="Zero-Inflated Poisson versus Negative Binomial Models",
          align=TRUE,
          column.labels=c("ZIP", "ZINB"),
          single.row=T,
          model.numbers=F)


```


```{r modComp, echo=FALSE,warning=FALSE,error=FALSE}

#windows(width=8,height=6,record=T)
resids=residuals(zipmod)
#cbind(fitted(zipmod),dat$SumCount,fitted(zipmod)-dat$SumCount,resids)
plot(fitted(zipmod),resids)
plot(dat$y,resids,xlab="Year",main="Residuals (zipmod)")
plot(dat$wc,resids,xlab="Water Clarity",main="Residuals (zipmod)")
plot(dat$cd,resids,xlab="Current Direction",main="Residuals (zipmod)")
plot(dat$bd,resids,xlab="Biotic Diversity",main="Residuals (zipmod)")
plot(dat$lat,resids,xlab="Latitude",main="Residuals (zipmod)")

plot(dat$SumCount,fitted(zipmod))

hist(dat$SumCount,breaks=0:max(dat$SumCount),freq=T,right=TRUE,xlab='Aggregate Fish Counted', main='ZIP')  
d=hist(predict(zipmod),breaks=0:max(dat$SumCount),plot=FALSE)
lines(seq(0.5,max(dat$SumCount),by=1),d$counts, col="blue",type='l')      
hist(dat$SumCount,breaks=0:max(dat$SumCount),freq=T,right=TRUE,xlab='Aggregate Fish Counted', main='ZIP',ylim=c(0,50))  
lines(seq(0.5,max(dat$SumCount),by=1),d$counts, col="blue",type='b')      


#windows(width=8,height=6,record=T)
resids=residuals(nbmod)
#cbind(fitted(nbmod),dat$SumCount,fitted(nbmod)-dat$SumCount,resids)
plot(fitted(nbmod),resids)
plot(dat$SumCount,fitted(nbmod))

plot(dat$y,resids,xlab="Year",main="Residuals (nbmod)")
plot(dat$wc,resids,xlab="Water Clarity",main="Residuals (nbmod)")
plot(dat$cd,resids,xlab="Current Direction",main="Residuals (nbmod)")
plot(dat$bd,resids,xlab="Biotic Diversity",main="Residuals (nbmod)")
plot(dat$lat,resids,xlab="Latitude",main="Residuals (nbmod)")

hist(dat$SumCount,breaks=0:max(dat$SumCount),freq=T,right=TRUE,xlab='Aggregate Fish Counted', main='NB')  
d2=hist(predict(nbmod),breaks=0:max(dat$SumCount),plot=FALSE)
lines(seq(0.5,max(dat$SumCount),by=1),d2$counts, col="blue",type='l')      
hist(dat$SumCount,breaks=0:max(dat$SumCount),freq=T,right=TRUE,xlab='Aggregate Fish Counted', main='NB',ylim=c(0,50))  
lines(seq(0.5,max(dat$SumCount),by=1),d2$counts, col="blue",type='b')      

plot(dat$SumCount,fitted(nbmod))
points(dat$SumCount,fitted(zipmod),col='red',pch=19)

par(mfrow=c(1,2))
hist(dat$SumCount,breaks=0:max(dat$SumCount),freq=T,right=TRUE,xlab='SumCount', main='ZIP',ylim=c(0,50), xlim=c(0,60))  
d=hist(predict(zipmod),breaks=0:max(dat$SumCount),plot=FALSE)
lines(seq(0.5,max(dat$SumCount),by=1),d$counts, col="blue",type='b') 
  
hist(dat$SumCount,breaks=0:max(dat$SumCount),freq=T,right=TRUE,xlab='SumCount', main='ZINB',ylim=c(0,50),xlim=c(0,60))  
d2=hist(predict(nbmod),breaks=0:max(dat$SumCount),plot=FALSE)
lines(seq(0.5,max(dat$SumCount),by=1),d2$counts, col="blue",type='b')   


```


**So this bit allows variable selection within the ZINB**

```{r ZINB VarSel, echo=FALSE,message=T, warning=T, cache=TRUE}
# NULL MODEL FORM
#nbform=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp )

###Remove water clarity and current variables from mean
nbform1=formula(SumCount~ y  + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp )
nbmod1=zeroinfl(nbform1,  dist = "negbin", link = "logit",data=dat);summary(nbmod1)

nbform2=formula(SumCount~ y  + wc + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp )
nbmod2=zeroinfl(nbform2,  dist = "negbin", link = "logit",data=dat);summary(nbmod2)

###Remove benthic variables from mean
nbform3=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp )
nbmod3=zeroinfl(nbform3,  dist = "negbin", link = "logit",data=dat);summary(nbmod3)

nbform4=formula(SumCount~ y + wc + cd + sc + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp )
nbmod4=zeroinfl(nbform4,  dist = "negbin", link = "logit",data=dat);summary(nbmod4)

###Remove depth from mean
nbform5=formula(SumCount~ y + wc + cd + sc + bd + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp)
nbmod5=zeroinfl(nbform5,  dist = "negbin", link = "logit",data=dat);summary(nbmod5)

#### Remove Season From mean 
nbform6=formula(SumCount~ y + wc + cd + sc + bd + d + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp)
nbmod6=zeroinfl(nbform6,  dist = "negbin", link = "logit",data=dat);summary(nbmod6)

#### Remove Latitude From mean 
nbform7=formula(SumCount~ y + wc + cd + sc + bd + d + t + temp |y + wc + cd + sc + bd + d + t + lat + temp)
nbmod7=zeroinfl(nbform7,  dist = "negbin", link = "logit",data=dat);summary(nbmod7)

#### Remove Temperature From mean 
nbform8=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat  |y + wc + cd + sc + bd + d + t + lat + temp)
nbmod8=zeroinfl(nbform8,  dist = "negbin", link = "logit",data=dat);summary(nbmod8)

#### Remove year From mean 
nbform9=formula(SumCount~ wc + cd + sc + bd + d + t + lat + temp  |y + wc + cd + sc + bd + d + t + lat + temp)
nbmod9=zeroinfl(nbform9,  dist = "negbin", link = "logit",data=dat);summary(nbmod9)

lrtest(nbmod1,nbmod)
lrtest(nbmod2,nbmod)
lrtest(nbmod3,nbmod)
lrtest(nbmod4,nbmod)
lrtest(nbmod5,nbmod)
lrtest(nbmod6,nbmod)
lrtest(nbmod7,nbmod)
lrtest(nbmod8,nbmod)
lrtest(nbmod9,nbmod)

AIC(nbmod,nbmod1,nbmod2,nbmod3,nbmod4,
    nbmod5,nbmod6,nbmod7,nbmod8,nbmod9)

AIC(nbmod)-AIC(nbmod1)
AIC(nbmod)-AIC(nbmod2)
AIC(nbmod)-AIC(nbmod3)
AIC(nbmod)-AIC(nbmod4)
AIC(nbmod)-AIC(nbmod5)
AIC(nbmod)-AIC(nbmod6)
AIC(nbmod)-AIC(nbmod7)
AIC(nbmod)-AIC(nbmod8)


## SECOND ROUND ## --> MOVING TO LOGISTIC SIDE
# nbmod is the best performing so start with that

# Remove wc from count
nbform1=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + cd + sc + bd + d + t + lat + temp )
nbmod1=zeroinfl(nbform1,  dist = "negbin", link = "logit",data=dat);summary(nbmod1)
# Remove cd from count
nbform2=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + sc + bd + d + t + lat + temp )
nbmod2=zeroinfl(nbform2,  dist = "negbin", link = "logit",data=dat);summary(nbmod2)
# Remove sc from count
nbform3=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + bd + d + t + lat + temp )
nbmod3=zeroinfl(nbform3,  dist = "negbin", link = "logit",data=dat);summary(nbmod3)
# Remove bd from count
nbform4=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + d + t + lat + temp )
nbmod4=zeroinfl(nbform4,  dist = "negbin", link = "logit",data=dat);summary(nbmod4)
# Remove d from count
nbform5=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + t + lat + temp )
nbmod5=zeroinfl(nbform5,  dist = "negbin", link = "logit",data=dat);summary(nbmod5)
# Remove t from count
nbform6=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + lat + temp )
nbmod6=zeroinfl(nbform6,  dist = "negbin", link = "logit",data=dat);summary(nbmod6)
# Remove lat from count
nbform7=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t +  temp )
nbmod7=zeroinfl(nbform7,  dist = "negbin", link = "logit",data=dat);summary(nbmod7)
# Remove temp from count
nbform8=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat )
nbmod8=zeroinfl(nbform8,  dist = "negbin", link = "logit",data=dat);summary(nbmod8)
# Remove y from count
nbform9=formula(SumCount~ y + wc + cd + sc + bd + d + t + lat + temp | wc + cd + sc + bd + d + t + lat + temp )
nbmod9=zeroinfl(nbform9,  dist = "negbin", link = "logit",data=dat);summary(nbmod9)


lrtest(nbmod1,nbmod)
lrtest(nbmod2,nbmod)
lrtest(nbmod3,nbmod)
lrtest(nbmod4,nbmod)
lrtest(nbmod5,nbmod)
lrtest(nbmod6,nbmod)
lrtest(nbmod7,nbmod)
lrtest(nbmod8,nbmod)
lrtest(nbmod9,nbmod)

AIC(nbmod,nbmod1,nbmod2,nbmod3,nbmod4,
    nbmod5,nbmod6,nbmod7,nbmod8,nbmod9)

AIC(nbmod)-AIC(nbmod1)
AIC(nbmod)-AIC(nbmod2)
AIC(nbmod)-AIC(nbmod3)
AIC(nbmod)-AIC(nbmod4)
AIC(nbmod)-AIC(nbmod5)
AIC(nbmod)-AIC(nbmod6)
AIC(nbmod)-AIC(nbmod7)
AIC(nbmod)-AIC(nbmod8)





## nbmod3 needs no further removals from mean


# Model 3 is the best performing for the positive part of the model
#nbform3=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat + temp )
##Now conduct variable selection on the binomial part
### Remove year
nbform3log1=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | wc + cd + sc + bd + d + t + lat + temp )
nbmod16=zeroinfl(nbform3log1,  dist = "negbin", link = "logit",data=dat);summary(nbmod16)

###Remove wc from mean
nbform3log2=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y +  cd + sc + bd + d + t + lat + temp )
nbmod17=zeroinfl(nbform3log2,  dist = "negbin", link = "logit",data=dat);summary(nbmod17)

###Remove current direction  from mean
nbform3log3=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc +  sc + bd + d + t + lat + temp )
nbmod18=zeroinfl(nbform3log3,  dist = "negbin", link = "logit",data=dat);summary(nbmod18)

## Remove substrat cat
nbform3log4=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc + cd + bd + d + t + lat + temp )
nbmod19=zeroinfl(nbform3log4,  dist = "negbin", link = "logit",data=dat);summary(nbmod19)

###Remove benthic variables from mean
nbform3log5=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc + cd + sc + d + t + lat + temp )
nbmod20=zeroinfl(nbform3log5,  dist = "negbin", link = "logit",data=dat);summary(nbmod20)

## Remove depth
nbform3log6=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc + cd + sc + bd + t  + lat + temp )
nbmod21=zeroinfl(nbform3log6,  dist = "negbin", link = "logit",data=dat);summary(nbmod21)

###Remove julian day
nbform3log7=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + lat  + temp )
nbmod22=zeroinfl(nbform3log7,  dist = "negbin", link = "logit",data=dat);summary(nbmod22)

#### Remove Latitude 
nbform3log8=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + temp  )
nbmod23=zeroinfl(nbform3log8,  dist = "negbin", link = "logit",data=dat);summary(nbmod23)

## Remove temp
nbform3log9=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp |y + wc + cd + sc + bd + d + t + lat  )
nbmod24=zeroinfl(nbform3log9,  dist = "negbin", link = "logit",data=dat);summary(nbmod24)




lrtest(nbmod16,nbmod3)
lrtest(nbmod17,nbmod3)
lrtest(nbmod18,nbmod3)
lrtest(nbmod19,nbmod3)
lrtest(nbmod20,nbmod3)
lrtest(nbmod21,nbmod3)
lrtest(nbmod22,nbmod3)
lrtest(nbmod23,nbmod3)
lrtest(nbmod24,nbmod3)
AIC(nbmod16)
AIC(nbmod17)
AIC(nbmod18)
AIC(nbmod19)
AIC(nbmod20)
AIC(nbmod21)
AIC(nbmod22)
AIC(nbmod23)
AIC(nbmod24)
AIC(nbmod3)-AIC(nbmod16)
AIC(nbmod3)-AIC(nbmod17)
AIC(nbmod3)-AIC(nbmod18)
AIC(nbmod3)-AIC(nbmod19)
AIC(nbmod3)-AIC(nbmod20)
AIC(nbmod3)-AIC(nbmod21)
AIC(nbmod3)-AIC(nbmod22)
AIC(nbmod3)-AIC(nbmod23)
AIC(nbmod3)-AIC(nbmod24)

## nmmod16 (drop year from binomial) has the lowest AIC and highest sig lrtest so 
## start with that and try dropping others in the binomial



# Model 2 is the best performing for the binomial part of the model
###Remove year from mean                                              
nbform3log10=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | cd + sc + bd + d + t + lat + temp)
nbmod25=zeroinfl(nbform3log10,  dist = "negbin", link = "logit",data=dat);summary(nbmod25)

###Remove current variables from mean
nbform3log11=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | wc + sc + bd + d + t + lat + temp)
nbmod26=zeroinfl(nbform3log11,  dist = "negbin", link = "logit",data=dat);summary(nbmod26)

###Remove benthic variables from mean
nbform3log12=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | wc + cd + bd + d + t + lat + temp)
nbmod27=zeroinfl(nbform3log12,  dist = "negbin", link = "logit",data=dat);summary(nbmod27)

nbform3log13=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | wc + cd + sc + d + t + lat + temp)
nbmod28=zeroinfl(nbform3log13,  dist = "negbin", link = "logit",data=dat);summary(nbmod28)

###Remove depth from mean
nbform3log14=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | wc + cd + sc + bd + t + lat + temp)
nbmod29=zeroinfl(nbform3log14,  dist = "negbin", link = "logit",data=dat);summary(nbmod29)

#### Remove Latitude From mean 
nbform3log15=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | wc + cd + sc + bd + d + lat + temp)
nbmod30=zeroinfl(nbform3log15,  dist = "negbin", link = "logit",data=dat);summary(nbmod30)

#### Remove Temperature From mean 
nbform3log16=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | wc + cd + sc + bd + d + t + temp)
nbmod31=zeroinfl(nbform3log16,  dist = "negbin", link = "logit",data=dat);summary(nbmod31)

#### Remove Temperature From mean 
nbform3log17=formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | wc + cd + sc + bd + d + t + lat )
nbmod32=zeroinfl(nbform3log17,  dist = "negbin", link = "logit",data=dat);summary(nbmod32)


lrtest(nbmod25,nbmod16)
lrtest(nbmod26,nbmod16)
lrtest(nbmod27,nbmod16)
lrtest(nbmod28,nbmod16)
lrtest(nbmod29,nbmod16)
lrtest(nbmod30,nbmod16)
lrtest(nbmod31,nbmod16)
lrtest(nbmod32,nbmod16)
AIC(nbmod25)
AIC(nbmod26)
AIC(nbmod27)
AIC(nbmod28)
AIC(nbmod29)
AIC(nbmod30)
AIC(nbmod31)
AIC(nbmod32)
AIC(nbmod16)-AIC(nbmod25)
AIC(nbmod16)-AIC(nbmod26)
AIC(nbmod16)-AIC(nbmod27)
AIC(nbmod16)-AIC(nbmod28)
AIC(nbmod16)-AIC(nbmod29)
AIC(nbmod16)-AIC(nbmod30)
AIC(nbmod16)-AIC(nbmod31)
AIC(nbmod16)-AIC(nbmod32)


###Best Model is nbmod25
# SumCount~ y + wc + cd + bd + d + t + lat + temp | cd + sc + bd + d + t + lat + temp

nbbest=nbmod25

```

The final model:
nbmod25

```{r ResidFitted, echo=FALSE, error=FALSE, warning=FALSE}

#windows(width=8,height=6,record=T)
resids=residuals(nbbest, type="pearson")

#cbind(fitted(nbbest),dat$SumCount,fitted(nbbest)-dat$SumCount,resids)
par(mfrow=c(1,2))
plot(fitted(nbbest),resids, ylab="Pearsons Residuals", xlab="Fitted Values") #pearson resids vs fitted values
plot(dat$SumCount,fitted(nbbest), ylab="Fitted Values", xlab="Original Values") #pearson values vs original data
```


```{r residVar, echo=FALSE, warning=FALSE, error=FALSE, fig.align='center'}

# A single figure
par(mfrow=c(2,2))
plot(dat$y,resids,xlab="Year",main="Residuals (nbbest)")
plot(dat$t,resids,xlab="Season",main="Residuals (nbbest)")
plot(dat$lat,resids,xlab="Latitude",main="Residuals (nbbest)")

par(mfrow=c(3,2))
plot(dat$wc,resids,xlab="Water Clarity",main="Residuals (nbbest)")
plot(dat$cd,resids,xlab="Current Direction",main="Residuals (nbbest)")
plot(dat$sc,resids,xlab="Substrate Composition",main="Residuals (nbbest)")
plot(dat$bd,resids,xlab="Biotic Diversity",main="Residuals (nbbest)")
plot(dat$d,resids,xlab="Depth",main="Residuals (nbbest)")
plot(dat$temp,resids,xlab="Temperature",main="Residuals (nbbest)")


```


```{r zinbFit, echo=FALSE, warning=FALSE, error=FALSE, fig.align='center'}
# Model Fit figure
par(mfrow=c(1,2))
hist(dat$SumCount,
     breaks=0:max(dat$SumCount),
     freq=T,right=TRUE,
     xlab='Sum Count',
     col ="lightgray",
     bor= "lightgray",
     main='NB')  
d3=hist(predict(nbbest),breaks=0:max(dat$SumCount),plot=FALSE)
lines(seq(0.5,max(dat$SumCount),by=1),d3$counts, col="blue",type='l')      
hist(dat$SumCount,
     breaks=0:max(dat$SumCount),
     freq=T,right=TRUE,
     xlab='Sum Count', 
     col = "gray",
     bor="gray",
     main='NB',
     ylim=c(0,50), xlim=c(0, 100))  
lines(seq(0.5,max(dat$SumCount),by=1),d3$counts, col="blue",type='l')      
#lines(seq(0.5,max(dat$SumCount),by=1),d2$counts, col="red",type='l')      
```


```{r idkPlot, echo=FALSE, warning=FALSE, error=FALSE, fig.align='center'}

#formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | cd + sc + bd + d + t + lat + temp)
new.dat=expand.grid(y=levels(dat$y),
                    wc=levels(dat$wc),
                    #cm=levels(dat$cm),
                    cd=levels(dat$cd),
                    sc=levels(dat$sc),
                    bd=levels(dat$bd),
                    #bt=levels(dat$bt),
                    #bh=levels(dat$bh),
                    d=levels(dat$d),
                    t=levels(dat$t),
                    lat=levels(dat$lat),
                    temp=levels(dat$temp)
                    #tod=levels(dat$tod),
                    #frames=1)
                    )

new.dat=cbind(new.dat,predict(nbbest,new.dat))
names(new.dat)[dim(new.dat)[2]]="Predicted"
resvec=summaryBy(Predicted~y,data=new.dat,FUN=mean)[,2]
index = resvec/mean(resvec)
#index = resvec/max(resvec)

plot(resvec,type='b',ylim=c(0,6))
plot(index,type='b',ylim=c(0,2))
```



```{r bootStrap, cache=TRUE, echo=FALSE, error=FALSE, warning=FALSE}

#set up data objects and specify number of bootstrap replications
ptm <- proc.time()
boots=5   #10000
##formula(SumCount~ y + wc + cd + bd + d + t + lat + temp | cd + sc + bd + d + t + lat + temp)

org.dat=dat[,c(1,3,4,5,7,10,13,15,16)];head(org.dat)
boot.dat=org.dat
numyrs=length(levels(org.dat$y));numyrs
index.boot=matrix(NA,nrow=boots,ncol=numyrs)
predmean.boot=index.boot
yr.samp=summaryBy(SumCount~y,data=org.dat,FUN=length)[,2] #added a left side to Lew's code base

#start bootstrap loop
for(boot in 1:boots){
  
  #Get bootstrap data for current replicate
  for(i in 1:numyrs){
    yr.rows=org.dat$y==levels(org.dat$y)[i]
    sub.dat=org.dat[yr.rows,]
    yr.boot=round(runif(yr.samp[i],1,yr.samp[i]),0)
    boot.dat[yr.rows,]=sub.dat[yr.boot,]
  }
  
# Now Recalibrate 2010 Data 
  cal.prop=0.61
  table(boot.dat$y,boot.dat$SumCount)
  dat.2010=boot.dat[boot.dat$y==2010,]
  cal.SumCount=rep(NA,length(dat.2010$SumCount))
  for(i in 1:length(dat.2010$SumCount)){cal.SumCount[i]=rbinom(1,dat.2010$SumCount[i],cal.prop)}#;plot(cal.SumCount,dat.2010$SumCount,xlim=c(0,250),ylim=c(0,250))
  boot.dat[boot.dat$y==2010,1]= cal.SumCount 
  
  
  #make a function to compute the index and return either a valid index or NA conditional on if the model converges
  getindex=function(){
    #define and fit model for current replicate.  Use the "try" function so that can continue to run if model does not converge
    #on a particular replication  
    f2 =formula(SumCount~ y + wc + cd + sc + bd + d + lat + temp |y + cd + sc + bd + d + lat + temp)
    Nb2 = try(zeroinfl(f2, dist = "negbin", link = "logit", data = boot.dat)); Nb2
    
    #see if model converged, if it did, return the mean year effect over all covariate combinations
    # If it did not converge, return an NA vector for the year effects
    if (class(Nb2) != "try-error"){
      #Predict the year effect (index) by predicting for each covariate level and compute the mean
    new.dat=expand.grid(y=levels(dat$y),
                    wc=levels(dat$wc),
                    #cm=levels(dat$cm),
                    cd=levels(dat$cd),
                    sc=levels(dat$sc),
                    bd=levels(dat$bd),
                    #bt=levels(dat$bt),
                    #bh=levels(dat$bh),
                    d=levels(dat$d),
                    t=levels(dat$t),
                    lat=levels(dat$lat),
                    temp=levels(dat$temp)
                    #tod=levels(dat$tod),
                    #frames=1)
                    )
      new.dat=cbind(new.dat,predict(Nb2,new.dat))
      names(new.dat)[10]="Predicted"
      resvec=summaryBy(Predicted~y,data=new.dat,FUN=mean)[,2]
      return(resvec)
    }  
    resvec=rep(NA,numyrs)
    return(resvec)
  }
  
  #Call the getindex function for each bootstrap replicate
  predmean.boot[boot,]=getindex()
  index.boot[boot,]=predmean.boot[boot,]/mean(predmean.boot[boot,])
}
save.image("\\\\CCFHR-S-1534090\\popdyn1\\Purcell\\GrayTrigIndex_SEDAR41\\GrayTrig_VideoIndex.RData")
```


```{r indexFig, echo=FALSE, warning=FALSE, error=FALSE}

#windows(width=8,height=6,record=T)
medianidex=apply(predmean.boot[!is.na(predmean.boot[,1]),],2,quantile,c(0.5))
plot(resvec)
lines(medianidex)
#exclude runs that did not converge and compute the convergence rate
conv.index=index.boot[!is.na(index.boot[,1]),]
convrate=dim(conv.index)[1]/dim(index.boot)[1]

convrate

#Calculate the CV by year

#Make a data frame from the boot strap output
cv.dat<-as.data.frame(index.boot)
cv.dat2<-cv.dat[complete.cases(cv.dat),]

# Make a function for coef vari
co.var<- function(x) (sd(x)/mean(x))

# Call co.var on each column of the bootstrap output df
cv.2011<-co.var(cv.dat2$V1)  
cv.2012<-co.var(cv.dat2$V2)
cv.2013<-co.var(cv.dat2$V3)


(proc.time() - ptm)[3]/60/60 #???
#conv.index

cI=apply(conv.index,2,quantile,c(0.025,0.5,0.975)) #97.5% CI
#cI=apply(conv.index,2,quantile,c(0.05,0.5,0.95))  #95% CI
nomcpue<-summaryBy(SumCount ~ y, dat=dat, FUN=mean)
nomcpue.std <- nomcpue$SumCount.mean/mean(nomcpue$SumCount)
nomcpue.std <- as.data.frame(nomcpue.std)


matplot(t(cI),type='l',ylim=c(0,2.25),xaxt="n",col=c(1,2,1),lty=c(3,1,3),lwd=c(1,2,1),ylab='Relative CPUE')
axis(1,at=1:3,labels=2011:2013)
index = resvec/mean(resvec)
#lines(index,type='l',ylim=c(0,2),col='green',lwd=2)
lines(nomcpue.std$nomcpue.std, type='l', col='blue', lty=2,lwd=2)
legend('topright',legend=c('Standardized Index ','Bootstrap CI','Nominal'),lty=c(1,3,2),col=c('red','black','blue'),lwd=c(2,1,2), cex=0.8, bty="n")

# yrs=as.numeric(row.names(ds$deltaGLM.index))
# errbar(x=yrs,y=U.std,yplus=U.std+1.96*error.std,yminus=U.std-1.96*error.std, ylab='Relative CPUE', xlab='Year', lwd=2, ylim=c(0,5))
# grid(col="gray")
# title(main="GOM Shrimp Abundance Index")
# lines(yrs,U.std,lwd=2)
# lines(yrs,nomcpue.std$nomcpue.std,lty=2, lwd=2)
# legend("topleft", bg="white", legend=c("Standardized index","Nominal"),lwd=2, lty=1:2,pch=c(1,-1))
```



```{r, sessionInfo, echo=FALSE, warning=FALSE, error=FALSE, results='hide'}
# produce R session info
sessionInfo()

save.image("\\\\CCFHR-S-1534090\\popdyn1\\Purcell\\GrayTrigIndex_SEDAR41\\GrayTrig_VideoIndex.RData")
```
